---
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
title: "AKSSF Temperature Data Summary"
output:
  html_document: 
    df_print: paged
    fig_width: 10
    fig_height: 6
    fig_caption: yes
    code_folding: hide
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: false
editor_options: 
  chunk_output_type: inline
---

Document last updated `r Sys.time()` by Rebecca Shaftel (rsshaftel@alaska.edu). 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)


# load packages
library(tidyverse)
library(lubridate)
library(readr)
library(readxl)
library(hms)
library(plotly)
library(DT)
library(leaflet)
library(sf)
library(tmap)
library(tmaptools)


```

This is a summary of the stream temperature dataset that we will are using the AKSSF Stream Thermal Sensitivity project.

We had a team meeting on January 25, 2021 and there were several things we discussed.

* We have included several additional datasets, including those recommended by Luca for the Copper River watershed and additional data from Kodiak.
* For daymet, air temperatures were extracted by site location. Daymet are on a 1 km grid. They have not been processed by catchment because the daymet netcdf subset service was down from mid-May until June 10 when I was working on this. Daily daymet air temperatures for catchments and sites are compared at ~60 locations where we had both in the Anchor and Kenai River watersheds, which is summarized below.
* There was some discussion during our first meeting regarding whether or how much of September to include in the analysis. I have created some data summaries showing the sample size after filtering on end dates of Aug. 31, Sept. 15, and  Sept. 30 and completeness values of 70%, 80%, and 90% (number of days within the summer window).
* Part of the September discussion included whether air temperatures were below zero towards the end of that month. I plotted the air-stream temperatures by month and region to visualize those ranges and patterns and also made a table looking at which days were below zero.
* Since DFA will be run by year, I summarized the number of sites by each year to see if there is a start year we should use or some years where data are too limited to include in the analysis.
* We also talked about an interest in exploring the response to 2019 in particular so I have created some plots showing data by region for that year.

Decisions to prepare data for the DFA.

1. Do we want the daymet air temperatures summarized over catchments or sites?
2. What should be use for our end date in the summer window?
3. How much missing data can DFA handle within a time series? Should we filter on 70%, 80%, or 90% completeness? 
4. Are there a specific set of years we want to focus on?
5. Anything else I am missing that should be considered before running the DFA?

# Map of sites and years of data

This map shows the location of all of the sites that we have data for. Selecting each marker will show the data provider, the site name, the river name, the start year, the end year, and the number of years of data. 

```{r read in data}

sumdat <- readRDS("data_preparation/summer_data_wair2021-06-16.rds")
md <- read_rds("data_preparation/final_data/md.rds")

md_sf <- st_as_sf(md, coords = c("Longitude", "Latitude"), crs = "wgs84")

map.dat <- sumdat %>% 
  distinct(SiteID, year = year(sampleDate)) %>%
  group_by(SiteID) %>% 
  summarize(start_year = min(year),
            end_year = max(year),
            total_years = n()) %>% 
  arrange(SiteID)

md_sf <- merge(md_sf, map.dat)
```


```{r leaflet map, eval = FALSE}

# create map
leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  #fitBounds(-150, 60.04,-149.0, 60.02) %>%
  #setView(-150.210169, 60.487694, zoom = 8) %>%
  addMarkers(lng = map.dat$Longitude, lat = map.dat$Latitude,
             popup = paste("SiteID = ", map.dat$SiteID, "<br>",
                           "Waterbody = ", map.dat$Waterbody_name, "<br>",
                           "Data Source = ", map.dat$SourceName, "<br>",
                           "Start Year = ", map.dat$startYear, "<br>",
                           "End Year = ", map.dat$endYear, "<br>",
                           "Total Years of Data = ", map.dat$totYears, "<br>"))

```


```{r tmap map}
tmap_mode("view")

sitemap <- md_sf %>%
  tm_shape() +
  tm_dots(id = "SiteID", size = 0.05, group = "Sites", col = "SourceName",
          popup.vars = c("Waterbody_name", "start_year", "end_year", "total_years")) +
  tm_text("SiteID", size = 1.5, shadow = TRUE, auto.placement = TRUE,
          just = "bottom", remove.overlap = TRUE, clustering = TRUE, group = "Labels" ) +
  tm_basemap(server = c(Topo = "Esri.WorldTopoMap", Imagery = "Esri.WorldImagery" ))

sitemap

```

# Compare site and catchment air temperatures

For a previous project, we averaged daymet air temperatures over catchments in the Anchor and Kenai River watersheds at a daily time step. These can be used to evaluate differences between site and catchment air temperatures. This is for 42 sites in the Anchor and 24 sites in the Kenai.

```{r}
air_comp <- readRDS("data_preparation/air_comp.rds")

aircor <- air_comp %>% summarize(cor(tair, airDT))  

air_comp %>% 
  ggplot() +
  geom_point(aes(x = tair, y = airDT)) +
  geom_text(aes(x = 5, y = 15, label = paste0("r = ", round(aircor, 2)))) +
  geom_abline(intercept = 0, slope = 1) +
  ylim(c(2,20)) +
  xlim(c(2,20)) +
  labs(x = "Mean Catchment Air Temperatures", y = "Site Air Temperatures", title = str_wrap("Correlation between air temperatures extracted by point versus averaged over the catchment for 66 sites in the Anchor and Kenai River watersheds", 60)) +
  theme_bw()

```


# Summer window and completeness

```{r}

#get huc8, region, and waterbody name on the data.

sumdat <- left_join(sumdat, md %>% select(SiteID, Waterbody_name, Name:Region))

sumtbl <- sumdat %>% 
  mutate(year = year(sampleDate),
         jd = format(sampleDate, "%j"),
         S30 = case_when(jd < 274 ~ 1,
                         TRUE ~ 0),
         S15 = case_when(jd < 259 ~ 1,
                         TRUE ~ 0),
         A31 = case_when(jd < 244 ~ 1,
                         TRUE ~ 0)) %>% 
  group_by(Region, SiteID, Waterbody_name, year) %>%
  summarize("Sept. 30" = sum(S30 == 1)/122,
            "Sept. 15" = sum(S15 == 1)/107,
            "Aug. 31" = sum(A31 == 1)/92) %>% 
  ungroup() %>% 
  pivot_longer(cols = 'Sept. 30':'Aug. 31', names_to = "Window", values_to = "value")
 
```

Table showing the number of time series (site and year combinations) with data completeness in each of three windows, all starting June 1. Window indicates the end date and the columns indicate the percent of days with data within the summer window (e.g. June 1 - September 30).

```{r}
sumtbl %>% 
  group_by(Window) %>% 
  summarize("70%" = sum(value > 0.7),
            "80%" = sum(value > 0.8),
            "90%" = sum(value > 0.9)) 


```




# Air-stream relationships by month

These are plots of the daily air temperatures and stream temperatures by month and region. There is obviously a lot of sites that have very cold stream temperatures with little response to air temperatures. There are some air temperatures below zero, especially in the Copper River watershed and also in Cook Inlet.

```{r}
sumdat %>% 
  mutate(month = month(sampleDate, label = TRUE)) %>% 
  ggplot() +
  geom_point(aes(y = meanDT, x = airDT)) +
  facet_grid(cols = vars(Region), rows = vars(month)) +
  theme_bw()
```


Table showing number of days with air temperatures below zero by month and day. Almost everything is after September 20, so cutting the data off at September 15 seems to make sense.

```{r}
sumdat %>% 
  filter(airDT < 0) %>% 
  mutate(day = format(sampleDate, "%m-%d")) %>% 
  count(day) %>% 
  arrange(day)
```

# Data summary by year

Function to filter data frame by different combinations of julian date and completeness. Note this assumes a lot about the input data frame.

```{r}

jd_completeness_filter <- function(inputDat, jd_filter, completeness_filter) {
  dat <- inputDat %>% 
    mutate(jd = as.numeric(format(sampleDate, "%j")),
           year = year(sampleDate)) %>% 
    filter(jd >= 152, jd <= jd_filter) #June 1st to end date
  ndays <- max(dat$jd) - min(dat$jd)
  dat2 <- dat %>%
    group_by(SiteID, year) %>%
    mutate(completeness = n()/ndays * 100) %>% 
    filter(completeness > completeness_filter) %>% 
    ungroup()
  return(dat2)
}

```

Filter on September 15 as an end date and 70% of days within summer window.

```{r}
sumdat2 <- jd_completeness_filter(sumdat, 258, 70)

sumdat2 %>% 
  distinct(SiteID, year) %>% 
  count(year) %>% 
  ggplot() +
  geom_point(aes(x = year, y = n)) +
  labs(x = "Year", y = "Number of Sites", title = "Sites with 70% of days in June 1 - September 15 window") +
  theme_bw()
```

# Patterns in 2019

Take the filtered data frame from last step and look at 2019.

```{r}
sumdat2 %>% 
  filter(year == 2019) %>% 
  complete(Region, SiteID, sampleDate = seq.Date(min(sampleDate), max(sampleDate), by = "day")) %>% 
  ggplot(aes(x = sampleDate, y = meanDT, group = SiteID)) +
  geom_line() +
  facet_wrap(~Region) +
  labs(x = "Date", y = "Mean Stream Temperature (°C)", title = "2019 Stream Temperatures by Region") +
  theme_bw()

```


