---
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
title: "AKSSF Temperature Data Summary"
output:
  html_document: 
    df_print: paged
    fig_width: 10
    fig_height: 6
    fig_caption: yes
    code_folding: hide
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: false
editor_options: 
  chunk_output_type: inline
---

Document last updated `r Sys.time()` by Rebecca Shaftel (rsshaftel@alaska.edu). 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)


# load packages
library(tidyverse)
library(lubridate)
library(readr)
library(readxl)
library(hms)
library(plotly)
library(DT)
library(leaflet)


```



# USFS Data 

Data from Luca Adelfio for 38 sites with 2 to 11 years (median = 7) of data from 2009 to 2019. Luca indicated that 11 sites were probably not useful in a landscape analysis because of site conditions and hydrology (e.g. lake influence, tidal influence, influence from downstream confluence).


```{r USFS site summary}
fsdat <- read_rds("output/fsdat.rds")

fsdat %>% 
  distinct(SiteID, year, useSite) %>%
  group_by(SiteID, useSite) %>% 
  summarize(start_year = min(year),
            end_year = max(year),
            count_yrs = n()) %>% 
  arrange(useSite) %>% 
  datatable()

```

This data summary table shows the mean, maximum, minimum, and sd of the mean daily temperatures for each site and year. These would be strongly influenced by the duration of the data. The number of daily observations within each year is also provided.


```{r USFS data summary}

fs.data.summary <- fsdat %>%
  group_by(SiteID, useSite, year, sampleDate) %>% 
  summarize(mndt = mean(Temperature)) %>% 
  group_by(SiteID, useSite, year) %>%
  summarize(meanTemp = mean(mndt, na.rm = T),
            maxTemp = max(mndt, na.rm = T),
            minTemp = min(mndt, na.rm = T),
            sdTemp = sd(mndt, na.rm = T),
            n_obs = n()) 

fs.data.summary %>%
  arrange(useSite) %>% 
  datatable() %>%
  formatRound(columns=c("meanTemp","maxTemp","minTemp","sdTemp"), digits=2)
```

This figures shows the mean daily temperatures for each site by year. There are some winter temps that we aren't interested in, but could be removed or flagged (e.g. < -1). Also, possibly some air temperatures in the open water season, check with Luca to see if these have been reviewed. (Note that raw subdaily temps show some pretty dramatic differences in daily ranges -- e.g. 25 mile, Bench Creek 2018, Chickaloon hws 2015 and 2016, Olsen Creek 2013.... maybe others?). 

```{r  USFS figure, fig.height=10, fig.width=8}

fsdat %>%
  filter(useSite == 1) %>% 
  group_by(SiteID, sampleDate) %>% 
  summarize(meanT = mean(Temperature)) %>%
  complete(SiteID, sampleDate = seq.Date(min(sampleDate), max(sampleDate), by = "day")) %>% 
  mutate(year = year(sampleDate),
         mo_day = format(sampleDate, "%m-%d")) %>% 
  ggplot(aes(x = as.Date(mo_day, format = "%m-%d"), y = meanT, color = as.factor(year))) +
  geom_line() +
  scale_x_date(date_breaks = "3 months", date_labels = "%b") +
  facet_wrap(~ SiteID) +
  labs(x = "Date", y = "Mean Daily Temperature", color = "Year",
       title = "USFS Original Logger Data by Site and Year") +
  theme_bw() +
  theme(legend.position = "bottom")
```

# Map of sites and years of data

This leaflet map shows the location of all of the sites that we have data for. Selecting each marker will show the data provider, the site name, the start year, the end year, and the number of years of data. Currently, only streams and rivers sites have been included.

```{r usfs}
usfs_md <- read_rds("output/usfs_md.rds")


fs.map.dat <- fsdat %>% 
  distinct(SiteID, year, useSite) %>%
  group_by(SiteID, useSite) %>% 
  summarize(startYear = min(year),
            endYear = max(year),
            totYears = n()) %>% 
  arrange(useSite) %>% 
  right_join(usfs_md)


fs.map.dat
```


```{r combine}
# all.dat <- bind_rows(uw.dat, accs.dat, cik.dat, fws.dat, nps.dat)
all.dat <- bind_rows(fs.map.dat)
```


```{r leaflet map}

# create map
leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  #fitBounds(-150, 60.04,-149.0, 60.02) %>%
  #setView(-150.210169, 60.487694, zoom = 8) %>%
  addMarkers(lng = all.dat$Longitude, lat = all.dat$Latitude,
             popup = paste("SiteID = ", all.dat$SiteID, "<br>",
                           "Data Source = ", all.dat$SourceName, "<br>",
                           "Start Year = ", all.dat$startYear, "<br>",
                           "End Year = ", all.dat$endYear, "<br>",
                           "Total Years of Data = ", all.dat$totYears, "<br>"))

```



