---
title: "CIK_trend_analysis"
output: 
  html_document:
    code_folding: hide
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 10, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = normalizePath(".."))

library(trend)
library(kableExtra)
library(knitr)
library(broom)
library(readxl)
library(stringr)
library(lubridate)
library(googlesheets4)
library(rnoaa)
library(hms)
library(tidyverse)


```

Source the function that assigns the mode of the time difference for each day to a formatted dataset. Also a function that calculates the number of days in a month.

```{r}
source("day_and_month_functions.R")
```

Load data.

```{r}
# cik.data <- readRDS("data_preparation/formatted_data/cik.data.rds")
# load cik.daily
```

# Temperature Figures

Create data frames of mean daily and mean monthly temperatures. Use functions that calculate the mode of the time difference for each day adn the number of days in a month. For screening 90% of measurements in a day and 90% of days in a month.
Filter on UseData == 1, which was added to the KNB dataset.

```{r}
cik.monthly <- cik.daily %>% 
  mutate(daysMonth = days_in_month(sampleDate),
         month = month(sampleDate),
         year = year(sampleDate)) %>% 
  group_by(SiteID, year, month, daysMonth) %>%
  summarize(meanMon = mean(meanDT),
            monCt = n()) %>% 
  filter(monCt > (0.9 * daysMonth))

# rm(cik.data)
```

Plot of daily means using complete so missing data don't have lines on ggplot.

```{r fig.height=11}
cik.daily %>%
  complete(SiteID, sampleDate = seq.Date(min(sampleDate), max(sampleDate), by = "day")) %>% 
  mutate(year = year(sampleDate),
         mo_day = format(sampleDate, "%m-%d")) %>% 
  ggplot(aes(x = as.Date(mo_day, format = "%m-%d"), y = meanT, color = as.factor(year))) +
  geom_line() +
  scale_x_date(date_breaks = "3 months", date_labels = "%b") +
  facet_wrap(~ SiteID) +
  labs(x = "Date", y = "Mean Daily Temperature", color = "Year",
       title = "CIK Original Logger Data by Site and Year") +
  theme_bw() +
  theme(legend.position = "bottom")
```

Same as above, but filtered on sites with 2019 data and that year is highlighted in red.

```{r fig.height=8}

sites_2019 <- cik.data %>% distinct(SiteID, year) %>% filter(year == 2019) %>% pull(SiteID)

cik.daily %>%
  filter(UseData == 1, SiteID %in% sites_2019) %>% 
  group_by(SiteID, Waterbody_name, sampleDate) %>% 
  summarize(meanT = mean(Temperature)) %>%
  complete(SiteID, sampleDate = seq.Date(min(sampleDate), max(sampleDate), by = "day")) %>% 
  mutate(year = year(sampleDate),
         mo_day = format(sampleDate, "%m-%d"),
         yr19 = case_when(year == 2019 ~ 1,
                          TRUE ~ 0)) %>% 
  ggplot(aes(x = as.Date(mo_day, format = "%m-%d"), y = meanT, color = as.factor(yr19), group = year)) +
  geom_line() +
  scale_color_manual(values = c("grey", "red"), labels = c("Other years", "2019")) +
  scale_x_date(date_breaks = "3 months", date_labels = "%b") +
  facet_wrap(~ Waterbody_name) +
  labs(x = "Date", y = "Mean Daily Temperature", color = "Year",
       title = "Stream Temperatures at CIK Sites with Data for 2019") +
  theme_bw() +
  theme(legend.position = "bottom", axis.title.x = element_blank())
```


Plot of monthly temperatures for sites with data for at least 9 years. There is a clear pattern of warm temperatures in 2003 - 2005 or so that makes linear trends hard to see. But, starting in 2006 through 2019, all streams seem to show a pretty clear pattern in increasing temperatures, especially in summer.

```{r fig.height=8}
ltSites <- cik.monthly %>% 
  ungroup() %>% 
  distinct(SiteID, year) %>% 
  count(SiteID) %>% 
  filter(n > 8) %>% 
  pull(SiteID)

cik.monthly %>% 
  filter(SiteID %in% ltSites) %>% 
  ggplot(aes(x = year, y = meanMon, color = as.factor(month))) + 
  geom_point() +
  # geom_line() +
  # geom_smooth(method = "lm") +
  facet_wrap(~SiteID) +
  theme_minimal()
```
All Sites.

```{r}
cik.monthly %>% 
  # filter(SiteID %in% ltSites) %>% 
  ggplot(aes(x = year, y = meanMon, color = as.factor(month))) + 
  geom_point() +
  # geom_line() +
  # geom_smooth(method = "lm") +
  facet_wrap(~Waterbody_name) +
  theme_minimal()
```




# Linear Trends in Monthly Temperatures

I decided looking at linear trends in mean monthly temperatures was the easiest analysis to pursue -- see Figure 4b) in Isaak DJ, Luce CH, Horan DL, et al (2018) Global Warming of Salmon and Trout Rivers in the Northwestern U.S.: Road to Ruin or Path Through Purgatory? Trans Am Fish Soc 147:566–587. https://doi.org/10.1002/tafs.10059. 

They also imputed missing data using a PCA, which might be really useful for the sites with some missing data -- e.g. Funny, Crooked, Slikok, and Soldotna. I'll look into this....

These are trends using any sites and months with more than 9 years of data.

```{r}
ltMonths <- cik.monthly %>% 
  ungroup() %>% 
  distinct(SiteID, year, month) %>% 
  count(SiteID, month) %>% 
  filter(n > 8)

monthlyLMs <- cik.monthly %>% 
  right_join(ltMonths) %>% 
  nest(data = c(-SiteID, -month)) %>% 
  mutate(
    fit = map(data, ~ lm(meanMon ~ year, data = .x)),
    tidied = map(fit, tidy)
  ) %>% 
  unnest(tidied) 

sig_trends <- monthlyLMs %>% 
  filter(p.value < 0.05, term == "year") %>% 
  select(month)

monthlyLMs %>% 
  filter(p.value < 0.05, term == "year") %>% 
  ggplot(aes(x = as.factor(month), y = estimate, color = SiteID)) +
  geom_point() +
  theme_minimal() +
  geom_abline(aes(intercept = 0, slope = 0)) 

```

Those estimates are really high, ~0.2 deg per year. Plot just these monthly trends and see if average monthly temperatures have actually increased by 2 deg over the last decade.

```{r}
cik.monthly %>% 
  right_join(sig_trends) %>% 
  ggplot(aes(x = year, y = meanMon, color = as.factor(month))) + 
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~Waterbody_name) +
  theme_minimal()
```

Just June and July trends, look to see if 2019 is really skewing those high.

```{r}
cik.monthly %>% 
  right_join(sig_trends) %>% 
  filter(month %in% 6:8) %>% 
  ggplot(aes(x = as.factor(year), y = meanMon, color = as.factor(month))) + 
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~Waterbody_name) +
  theme_minimal()
```

How much do trends decrease if we didn't include 2019? 

```{r fig.height=8}
monthlyLMs2 <- cik.monthly %>% 
  right_join(ltMonths) %>% 
  filter(year < 2019) %>% 
  nest(data = c(-SiteID, -month)) %>% 
  mutate(
    fit = map(data, ~ lm(meanMon ~ year, data = .x)),
    tidied = map(fit, tidy)
  ) %>% 
  unnest(tidied)

sig_trends2 <- monthlyLMs2 %>% 
  filter(p.value < 0.05, term == "year") %>% 
  select(SiteID, month)

nrow(sig_trends) - nrow(sig_trends2) 

monthlyLMs2 %>% 
  filter(p.value < 0.05, term == "year") %>% 
  ggplot(aes(x = as.factor(month), y = estimate * 10, color = SiteID)) +
  geom_point(size = 2) +
  theme_minimal() +
  geom_abline(aes(intercept = 0, slope = 0)) +
  labs(x = "Month", y = "Decadal Trend in Mean Monthly Temperature", title = "Significant Trends (p-value < 0.05) in Mean Monthly Temperatures \n from 2002-2018 for Cook Inlet Streams", color = "Stream: ") +
  theme(legend.position = "bottom")
```
Plot of significant trends without 2019.

```{r}
cik.monthly %>% 
  right_join(sig_trends2) %>% 
  # filter(month %in% 6:8) %>% 
  ggplot(aes(x = year, y = meanMon, color = Waterbody_name)) + 
  geom_jitter() +
  geom_smooth(method = "lm") +
  facet_wrap(~month) +
  theme_minimal()
```


Why does Ninilchik River trend for September drop when 2019 isn't included?

```{r}
cik.monthly %>% 
  filter(Waterbody_name == "Ninilchik River", month == 9) %>% 
  lm(meanMon ~ year, data = .) %>% 
  summary()

cik.monthly %>% 
  filter(Waterbody_name == "Ninilchik River", month == 9, year < 2019) %>% 
  lm(meanMon ~ year, data = .) %>% 
  summary()
```



Change in trends from entire time series (2002-2020) to time series with 2019 removed.

```{r}
bind_rows(
  monthlyLMs %>% 
    filter(p.value < 0.05, term == "year") %>% 
    mutate(trend = "w2019") %>% 
    select(Waterbody_name, estimate, trend, month),
  monthlyLMs2 %>% 
    filter(p.value < 0.05, term == "year") %>% 
    mutate(trend = "no2019") %>% 
    select(Waterbody_name, estimate, trend, month)
) %>% 
  pivot_wider(names_from = trend, values_from = estimate) %>% 
  kable(digits = 2)
  
```

Non-parametric trend test - Mann-Kendall followed by Sen Slope estimator.

```{r}

mk.list <- list()
sen.list <- list()

for(i in 1:nrow(ltMonths)) {
  dat <- cik.monthly %>% right_join(ltMonths %>% slice(i)) %>% pull(meanMon)
  mk.list[[i]] <- mk.test(dat)
  # names(mk.list[[i]]) <- paste(site_month[i, "SiteID"], site_month[i, "month"])
  sen.list[[i]] <- sens.slope(dat)
  # names(sen.list[[i]]) <- paste(site_month[i, "SiteID"], site_month[i, "month"])

}

mk.list
```

#Linear Trends in Seasonal Temperatures



# Missing data patterns

```{r}
all <- cik.monthly %>% right_join(ltMonths) %>% ungroup() %>% expand(SiteID, year, 6:9) %>% rename(month = `6:9`)

cik.monthly %>% 
  right_join(ltMonths) %>% 
  ungroup() %>% 
  count(SiteID)

cik.monthly %>% 
  ungroup() %>% 
  right_join(all) %>% 
  count(month, year, missing = is.na(meanMon)) %>%
  filter(missing == FALSE) %>% 
  ggplot(aes(x = year, y = month)) +
  geom_tile(aes(fill = n))
```

# Imputed data
Missing value imputation via missMDA. Following Josse and H J of Stat Software example staring on page 10.

I think I want to impute missing monthly means for sites by each year.

```{r}
library(missMDA)

#2009
sum2009 <- cik.monthly %>%
  ungroup() %>% 
  filter(year == 2009, month %in% 6:9) %>% 
  select(Waterbody_name, month, meanMon) %>% 
  pivot_wider(names_from = month, values_from = meanMon)

ncomp <- estim_ncpPCA(sum2009 %>% select(-Waterbody_name), ncp.min = 0, ncp.max = 6)
ncomp$ncp #1

res.imp <- imputePCA(sum2009 %>% select(-Waterbody_name), ncp = ncomp$ncp)

impute.df <- data.frame(bind_cols(Waterbody_name = sum2009 %>% pull(Waterbody_name),
                                  res.imp$completeObs)) %>%
  rename(`6` = X6, `7` = X7, `8` = X8, `9` = X9) %>% 
  mutate(impute = "imputed")

res.pca <- prcomp(res.imp$completeObs)
biplot(res.pca)

impute09 <- bind_rows(sum2009 %>% mutate(impute = "observed"), impute.df)

impute09 %>% 
  pivot_longer(cols = `6`:`9`, names_to = "month", values_to = "mean") %>%
  pivot_wider(names_from = impute, values_from = mean) #%>% 
  mutate(missing = case_when(is.na(observed) ~ "yes",
                             TRUE ~ "no")) %>% 
  ggplot(aes(x = month, y = imputed, color = missing)) +
  geom_jitter() +
  facet_wrap(~Waterbody_name)
          
```

Impute data for June through September for 14 long-term sites.

```{r}

sum.df <- cik.monthly %>%
  ungroup() %>% 
  filter(month %in% 6:9, SiteID %in% ltSites) %>% 
  mutate(mo_yr = paste(month, year, sep = "_")) %>% 
  select(Waterbody_name, mo_yr, meanMon) %>%
  pivot_wider(names_from = mo_yr, values_from = meanMon)

ncomp <- estim_ncpPCA(sum.df %>% select(-Waterbody_name), ncp.min = 0, ncp.max = 6)
ncomp$ncp #3

res.imp <- imputePCA(sum.df %>% select(-Waterbody_name), ncp = ncomp$ncp)

res.pca <- prcomp(res.imp$completeObs)
biplot(res.pca)

impute.df <- data.frame(bind_cols(Waterbody_name = sum.df %>% pull(Waterbody_name),
                                  res.imp$completeObs)) %>%
  mutate(impute = "imputed") %>% 
  pivot_longer(cols = X6_2008:X9_2019, names_to = "month", values_to = "mean") %>% 
  mutate(month = gsub("X", "", month))

impute.all <- sum.df %>% 
  mutate(impute = "observed") %>% 
  pivot_longer(cols = `6_2008`:`9_2019`, names_to = "month", values_to = "mean") %>% 
  bind_rows(impute.df)

impute.all %>% 
  pivot_wider(names_from = impute, values_from = mean) %>% 
  mutate(missing = case_when(is.na(observed) ~ "yes",
                             TRUE ~ "no")) %>% 
  ggplot(aes(x = month, y = imputed, color = missing)) +
  geom_jitter() +
  facet_wrap(~Waterbody_name)

  
```

Evaluate trends in monthly means on imputed values. Strange results, probably the matrix only has 14 sites and many month-year combinations only one or two sites had data. I think this is really only useful for imputing values for months in years where most sites were online, but maybe a logger was lost at a site or two.

```{r}

impute.df %>%
  rename(mo_yr = month) %>% 
  mutate(month = substr(mo_yr, 1, 1),
         year = as.numeric(substr(mo_yr, 3, 7))) %>% 
  ggplot(aes(x = year, y = mean, color = month)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~Waterbody_name)

monthlyLMs3 <- impute.df %>%
  rename(mo_yr = month) %>% 
  mutate(month = substr(mo_yr, 1, 1),
         year = as.numeric(substr(mo_yr, 3, 7))) %>% 
  filter(year < 2019) %>% 
  nest(data = c(-Waterbody_name, -month)) %>% 
  mutate(
    fit = map(data, ~ lm(mean ~ year, data = .x)),
    tidied = map(fit, tidy)
  ) %>% 
  unnest(tidied)

sig_trends3 <- monthlyLMs3 %>% 
  filter(p.value < 0.05, term == "year") #%>% 
  select(Waterbody_name, month)

monthlyLMs3 %>% 
  filter(p.value < 0.05, term == "year") %>% 
  ggplot(aes(x = as.factor(month), y = estimate, color = Waterbody_name)) +
  geom_point(size = 2) +
  theme_minimal() +
  geom_abline(aes(intercept = 0, slope = 0)) +
  labs(x = "Month", y = "Annual Trend in Mean Monthly Temperature", title = "Significant Trends (p-value < 0.05) in Mean Monthly Temperatures \n from 2002-2018 for Cook Inlet Streams", color = "Stream: ") +
  theme(legend.position = "bottom")
```

