---
title: "data_KNB"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath(".."))


library(tidyverse)
library(lubridate)
library(rnoaa)
library(plotly)
```


# Cook Inletkeeper

Sue's data archived on KNB are included in the data_CIK script.


# NPS - Trey Simmons

NPS data on KNB archive were moved to the data_NPS report so everything is in one place for this data source.

# ADFG - INCOMPLETE DATA 

## Read in data

Read in data files saved on S drive. There are four stream and four lakes sites. We only want streams for this project. For some reason, three are flagged as discrete data, but they are all hourly so import all four.

```{r}
adfg.files <- list.files("S:/Stream Temperature Data/ADFG Kodiak/data", pattern = ".csv", full.names = TRUE)

adfg.files <- adfg.files[!grepl("SiteLevelMetadata_Kodiak", adfg.files)]

```

Metadata

```{r}
adfg_md <- read_csv("S:\\Stream Temperature Data\\ADFG Kodiak\\data\\SiteLevelMetadata_Kodiak.csv") 

adfg_akssf_md <- adfg_md %>% 
  filter(Waterbody_type == "S") %>% 
  select(AKOATS_ID, SiteID, SourceName, Contact_person, Latitude, Longitude, Waterbody_name)

adfg_akssf_md

```

Combined data files and get information from metadata file.

```{r}
adfg.data <- adfg.files %>% 
  map_df(function(x) read_csv(x) %>% 
  mutate(file_name = gsub(".csv","",basename(x))))

adfg.data <- left_join(adfg.data, adfg_akssf_md %>% select(AKOATS_ID, SiteID, Waterbody_name)) %>% 
  filter(!is.na(SiteID)) %>% 
  mutate(year = year(sampleDate),
         dt = parse_datetime(paste(sampleDate, sampleTime), format = "%Y-%m-%d %H:%M:%S")) 

adfg.data %>% distinct(SiteID)

adfg.data %>% 
  group_by(SiteID) %>% 
  summarize(min(sampleDate),
            max(sampleDate))

```


## Review data

Plot of raw data. These data started in August 2016 and ended mid-summer 2017, not really useable for our analysis.

```{r}
adfg.data %>% count(UseData)

adfg.sites <- adfg.data %>% distinct(SiteID, year)

pdf("output/ADFG raw data by site.pdf")

for(i in 1:nrow(adfg.sites)) {
  dat <- left_join(adfg.sites %>% slice(i), adfg.data)
  p1 <- dat %>% 
    # filter(UseData == 1) %>%
    ggplot(aes(x = dt, y = Temperature)) +
    geom_line() +
    facet_wrap(~year) +
    labs(title = dat %>% distinct(SiteID))
  print(p1)  
}

dev.off()


```

Get air temperature data from a nearby GHCN site: GHCND:USR0000AKLA, using Dustin's code from Bristol Bay Temp repo. Bad data in first year, filter to 2000 on.

```{r}

air.dat <- meteo_pull_monitors("USR0000AFPK")  

air.dat <- air.dat %>% 
  # Temperature and Precipitation values are in tenths of degree/mm
  mutate_if( is.numeric, ~ . * 0.1) %>% 
  mutate(year = year(date)) 

summary(air.dat)

air.plot <- air.dat %>% 
  ggplot( aes( x = date)) +
  geom_line( aes(y = tmin, color = "Air min")) +
  geom_line( aes(y = tmax, color = "Air max")) 

ggplotly(air.plot) 

```

QA notes:

* all sites started in August 2016 so 2016 data are out. Just focus on June - Sept 2017.

* kdk_ayarv01 may have some air temps in early 2017, but June through August 2017 data look fine.
* kdk_doscr01 - same as previous site.
* kdk_frafp01a looks fine, the air temperature site doesn't seem very representative for Kodiak.
* kdk_karrv01 - drop this site because data collection ended in May 2017.


```{r}

adfg.daily <- adfg.data %>% 
  group_by(SiteID, sampleDate) %>% 
  summarize(meanTemp = mean(Temperature),
            n = n()) %>%
  filter(n %in% c(24, 48))

adfg.daily %>% distinct(SiteID)

p <- ggplot() +
  geom_line(data = air.dat, aes(x = date, y = tmin, color = "blue")) +
  geom_line(data = air.dat ,aes(x = date, y = tmax, color = "red")) +
  geom_line(data = adfg.daily %>% filter(SiteID == "kdk_karrv01"),
            aes(x = sampleDate, y = meanTemp))
ggplotly(p)
```

```{r}
adfg.data <- adfg.data %>% 
  filter(!SiteID == "kdk_karrv01", sampleDate > as.Date("2017-05-31"))
```


# Combine as daily mean temps and save

```{r}

knb.daily <- bind_rows(cik.data %>% filter(UseData == 1), nps.data, adfg.data) %>% 
  group_by(SiteID, Waterbody_name, year, sampleDate) %>% 
  summarize(meanTemp = mean(Temperature))


```


