---
title: "UWFWS data for Cook Inlet, Prince William Sound, Copper River and Kodiak"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE, messages = FALSE)
knitr::opts_knit$set(root.dir = normalizePath("..")) #this sets the root.dir up one level back to the project so that paths are relative to the project directory.

library(readxl)
library(stringr)
library(lubridate)
library(googlesheets4)
library(rnoaa)
library(hms)
library(tidyverse)
```



# Define Functions
Define any functions that are used for data formatting or duplicate measurement identification



```{r Functions}
source("W:/Github/AKSSF/helper_functions.R")

```

# Read in data and format


## Metadata

Sites file. This is a longer list of names from akoats that we should probably keep. Some basic ones should be filled in for providers with data not entered into akoats.

Notes on new names:

* Agency_ID should equal the SiteID in the data table. So during left_join, use by = c("SiteID" = "Agency_ID"). If this is a problem because a data provided used stream names, concatenate the agency acronym to the Agency_ID. E.g. USFS_Cold Creek.
* AKOATS_ID = seq_id.


```{r}


```



## Data

# Review data


## Duplicate measurements

?? Sufficient to just use distinct() on site/temp/dt or more robust method?

## Save data 
Save copies of metadata file and temp data formatted for QA to local drive
Save copies of metadata file and temp data formatted for QA to google drive

```{r Save Outputs}
# Save copy of formatted data for qc (UseSite vs UseData?)
# Save metadata - this will save a local copy and upload to google drive
save_metadata_files(data.in, acronym = "NAME OF OUTPUT")


```


If data have been reviewed by the data provider, these should be .csv of the final data, metadata, and daily data (see data_QA_TEMPLATE). Otherwise, an .rds or .csv to be passed to the data QA script.
