---
title: "data_USFS_Chugach"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath(".."))

library(tidyverse)
library(readxl)
library(stringr)
library(lubridate)
library(googlesheets4)
library(rnoaa)
library(hms)
#library(googledrive) #Only need if accessing file stored in drive to download

```


```{r Define Functions}

# Read in csv and add column with filename
read_csv_and_name <- function(csv_file_path) {
  sheet_name <- str_match(csv_file_path, "\\/\\s*(.*?)\\s*\\.csv")[2]
  dat <- read_csv(csv_file_path) %>% 
      mutate(file_name = sheet_name)
}

# Function to pull logger serial# from temp column
log_sn_fun <- function(string){ 
  
  str_extract(string, "\\-*\\d+\\.*\\d*")
} 

```


# Get air temperatures from GHCN sites using rnoaa

Not working, I can't find a site with daily summaries in cordova. The airport is part of the local climatological data online - WBAN:26410, but it's not working with the lcd function.

DM Notes: Can't get lcd function to work either but here is a workaround.

```{r NOAA data, message=FALSE, warning=FALSE}

#Token obtained from NOAA to access API
noaaTok <- "LmempjqpgcLSxDQWiLvuaOAGmscrQCrb"

# lcd(station = "26410", year = 2013)
# 
# ncdc_locs(locationcategoryid = "CITY")

#Station Codes for area of interest
cd.climStat <- c( "USC00502179","USW00026410", "USW00096405")


cd.climDat <- tibble( name = c( "CORDOVA WWTP", "CORDOVA AIRPORT",
"CORDOVA 14"),
                 id = cd.climStat)

# Pull Climate data from Cordova Airport
climDat <- meteo_pull_monitors(cd.climStat)  
str(climDat)
  

cd.climDat <- cd.climDat %>% 
  left_join( climDat[,c( "id", "date", "tmax", "tmin")], by = 'id') %>% 
  filter( date >= "2008-06-01",
          date <= "2020-12-30",
          name == "CORDOVA AIRPORT") %>% 
  # Temperature and Precipitation values are in tenths of degree/mm
  mutate_if( is.numeric, ~ . * 0.1) %>% 
  mutate(year = as.factor(year(date)),
         day = yday(date),
         dt = as.POSIXct(paste(date), format = "%Y-%m-%d"))
  

cd.climDat

```


# USFS
## Metadata

Priscilla was able to link most of these file names to site names in AKOATS. For the four sites not in AKOATS, she converted the UTM 
coordinates to latitude/longitude. Read in that google sheet so that we can get simple site names on the data sheet and also start a metadata file with the locations.

```{r read in metadata}

# Access google sheet with metadata for usfs chugach sites
depracated_metadata <- "https://docs.google.com/spreadsheets/d/1ymyhRyAomnJZTZqr4IbRmCnJjVaOIpCQYGscJDbCk-w/edit#gid=0"

usfs_dpmd <- read_sheet(depracated_metadata, sheet = "Sheet1",
                    col_names = TRUE,
                    col_types = "c") %>% 
  filter(ACCS_Project == "USFS_Chugach") %>% 
  mutate(AKOATS_ID = as.numeric(AKOATS_ID),
         Latitude = as.numeric(Latitude),
         Longitude = as.numeric(Longitude)) 

usfs_dpmd %>% select(Agency_ID, AKOATS_ID, Latitude, Longitude, data_SiteID)

```

Read in current copy of akoats so we can get lat longs for Luca's sites that Priscilla could match.

DM Notes: Use GOOGLE SHEET Copy for consistency

```{r akoats}

# Find and copy AKOATS working excel file in Google Drive
# x <- drive_find(pattern = "<NAME OF EXCEL FILE STORED IN GOOGLE DRIVE>", type = "xlsx", n_max = 1)
# akoats_cont <- drive_download(x, overwrite = TRUE)
# akoats_path <- akoats_cont$local_path
# akoats_fsch <- read_excel( akoats_path, sheet = "CONTINUOUS_DATA",
#                         col_names = TRUE
#                         )  %>%

# Access Google Sheet copy of AKOATS working sheet and use AKOATS Complete sheet

akoats_2020wk <- "https://docs.google.com/spreadsheets/d/1SPZXNGm_Tc39-GuJXY8j7Eb1lX6DwXTQ1en2_LvCI1I/edit#gid=1281874712"

akoats_fsch <- read_sheet(akoats_2020wk, sheet = "AKOATS_COMPLETE",
                    col_names = TRUE,
                    col_types = "c") %>% 
  select(seq_id,Agency_ID,Contact_person,SourceName,Contact_email,
         Contact_telephone,Latitude,Longitude,Sensor_accuracy) %>%
  rename(AKOATS_ID = seq_id) %>% 
         #SiteID = Agency_ID) %>% 
  mutate(AKOATS_ID = as.numeric(AKOATS_ID),
         Latitude = as.numeric(Latitude),
         Longitude = as.numeric(Longitude))

akoats_fsch <- akoats_fsch %>% 
  filter(SourceName == "usfsakChugach")
akoats_fsch


```
Should we create some AKOATS_IDs for the sites that do not have corresponding values?

```{r format fs sites, message=FALSE, warning=FALSE}

usfs_md <- tibble()
fs_akoats_sites <- tibble()

fs_akoats_sites <- usfs_dpmd %>% 
  select(Agency_ID, AKOATS_ID) %>% 
  left_join(akoats_fsch) %>% 
  filter(!is.na(AKOATS_ID))
  
keep <- intersect(names(fs_akoats_sites), names(usfs_dpmd))

usfs_md <- bind_rows(fs_akoats_sites,
  usfs_dpmd %>% filter(is.na(AKOATS_ID)) %>%
    select(one_of(keep)))

# saveRDS(usfs_md, "output/usfs_md.rds")

usfs_md

```

## Data

Luca provided a big chunk of data in ~2018 that was formatted consistently. He recently sent over some additional data for four sites from the Kenai from 2018-2020 that should be read in separately. 


```{r kenai new data, message=FALSE, warning=FALSE}

# # Locate Kenai 2018-2020 data
# fs_files2 <- list.files("C:\\Users\\dwmerrigan\\Documents\\GitHub\\AKSSF\\data_preparation\\source\\06_USFS_Chugach\\Kenai_2018-20", full.names = TRUE)
# 
# center_cr <- read_csv(fs_files2[grepl("CenterCrk", fs_files2)]) %>%
#   rename(Date = "Date Time,",
#          Time = "GMT-08:00",
#          Temperature = "Temp, °C (LGR S/N: 20163556, SEN S/N: 20163556)") %>%
#   mutate(Date = as.Date(Date, format = "%m/%d/%Y"),
#          Agency_ID = "Center Creek")
# 
# bench_cr <- read_excel(fs_files2[grepl("BENCH", fs_files2)]) %>%
#   rename(Date = "Date",
#          Time = "Time",
#          Temperature = "Temp, °C (LGR S/N: 10515702, SEN S/N: 10515702)") %>%
#   mutate(Date = as.Date(Date, format = "%m/%d/%Y"),
#          Time = as_hms(Time),
#          Agency_ID = "Bench Creek")
# 
# daves_cr <- read_excel(fs_files2[grepl("DavesCrk_surfaceH2O_cz_2018", fs_files2)]) %>%
#   rename(Date = "Date Time,",
#          Time = "GMT-08:00",
#          Temperature = "Temp, °C (LGR S/N: 10497829, SEN S/N: 10497829)") %>%
#   mutate(Date = as.Date(Date, format = "%m/%d/%Y"),
#          Time = as_hms(Time),
#          Agency_ID = "Daves Creek")
# 
# quartz_cr <- read_excel(fs_files2[grepl("QuartzCrk_surfaceH2O_cz_2018", fs_files2)]) %>%
#   rename(Date = "Date Time,",
#          Time = "GMT-09:00",
#          Temperature = "Temp, °C (LGR S/N: 20163535, SEN S/N: 20163535)") %>%
#   mutate(Date = as.Date(Date, format = "%m/%d/%Y"),
#          Time = as_hms(Time),
#          Agency_ID = "Quartz Creek")
# 
# kendat <- bind_rows(quartz_cr, daves_cr, bench_cr, center_cr)
# kendat

```

Extract file name of data set and in new column

```{r Collect temperature data files}

# Folder containing source datasets
source_fol <- "C:\\Users\\dwmerrigan\\Documents\\GitHub\\AKSSF\\data_preparation\\source"

files <- list.files(source_fol, full.names = T, recursive = T, pattern = ".*.csv|.*.xlsx")

# Remove metadata sheets and air temp datasets
patterns <- c("siteinfo", "selectedsites", "AirT", "geospatial", "UTM")

fs_files <- files
for (pattern in patterns){
  fs_files <- fs_files[!grepl(pattern, fs_files)]
}

basename(fs_files)
```

Bind all input datasets together after removing Airtemp and site info CSVs

```{r Merge data}

# Create tibble for fs temperature data
fsdat <- tibble()

# Create counters to check data file formats
t1 <- 0
t2 <- 0

for ( i in fs_files){
  # CSV Data
  if( endsWith (i,".csv")){
    dat <- read_csv( file = i)
    name = str_sub(basename(i), end = -5)
    
    if(grepl("/", dat[,1]) == TRUE) {
      dat <- dat[1:3]
      colnames(dat) <- c("sampleDate", "sampleTime", "Temperature")
      print( paste0( name, " has Date as first column"))
      print(colnames(dat))
      # Add to correctly formatted counter
      t1= t1 + 1
      
      # Format data
      dat <- dat %>% mutate(data_filename = name) %>%
        transform(sampleDate = as_date(mdy(sampleDate)),
                  sampleTime = hms::as.hms(sampleTime),
                  Temperature = as.numeric(Temperature))
      fsdat <- bind_rows(fsdat, dat[!is.na(dat$Temperature),]) %>%
        select(data_filename, sampleDate, sampleTime, Temperature)

    }else{
      print(paste0( name,  " has Different format "))
      # Add to incorrectly formatted counter
      t2= t2 + 1
    }
   
    # Excel Data  
  }else if( endsWith (i,".xlsx")){
    dat <- read_excel(path = i)
    name = str_sub(basename(i), end = -6)
    
    if(grepl("-", dat[,2]) == TRUE) {
      dat <- dat[1:3]
      colnames(dat) <- c("sampleDate", "sampleTime", "Temperature")
      print( paste0( name, " has Date as first column"))
      print(colnames(dat))
      # Add to correctly formatted counter
      t1= t1 + 1
      
      # Format data 
      dat <- dat %>% mutate(data_filename = name) %>% 
        transform(sampleDate = as.Date(sampleDate, format = "%m/%d/%Y"),
                  sampleTime = hms::as.hms(sampleTime),
                  Temperature = as.numeric(Temperature))
      fsdat <- bind_rows(fsdat, dat[!is.na(dat$Temperature),]) %>% 
        select(data_filename, sampleDate, sampleTime, Temperature)
      
    }else{
      print(paste0( name,  " has Different format "))
      # Add to incorrectly formatted counter
      t2= t2 + 1
    }
    
  }
}

print( paste0(t1, " Formatted correctly" ," | ",t2, " Incorrectly formatted"))
dput(unique(fsdat$data_filename))
```

```{r}
fs_site_names <- unique(fsdat$data_filename)
fs_site_names
 
as.data.frame(fs_site_names) %>% 
  write_csv(path = "data_preparation/formatted_data/usfs_site_names.csv")

```



```{r read in data from csv, message=FALSE, warning=FALSE}

# # Remove QAQC datasets
# fs_files <- fs_files[!grepl("QAQC", fs_files)] 

# Extract file name and save to new column for fs_files 
fsdat <- fs_files %>% 
  map_df(function(x) read_csv_and_name(x)) 

fsdat

```

Some problems with dates - some are 2 digit and some are 4 digit. In as.Date function, tryFormats won't fix this problem, which is unfortunate. I'll need to get number of digits using regexp and then manually tell it what format it is in.

```{r fix dates, message=FALSE, warning=FALSE}

fsdat %>% 
  mutate(year1 = sub(".*/.*/", "", Date)) %>% 
  distinct(year1)

fsdat <- fsdat %>% 
  mutate(year1 = sub(".*/.*/", "", Date),
         sampleDate = case_when(nchar(year1) == 4 ~ as.Date(Date, format = "%m/%d/%Y"),
                                TRUE ~ as.Date(Date, format = "%m/%d/%y"))) %>% 
  rename(sampleTime = Time, Temperature = Temp)

# fsdat %>%
#   group_by(file_name) %>%
#   summarize(min(sampleDate), max(sampleDate))
# 
# fsdat %>%
#   distinct(year = year(sampleDate)) %>%
#   arrange(year)
#
# fsdat %>%
#   distinct(file_name, year = year(sampleDate)) %>%
#   count(file_name) %>%
#   arrange(n) %>%
#   summarize(median(n))

fsdat

```

Additional fields to add to data frame:

* Add SiteIDs to the data. Unfortunately, Luca is using waterbody names for site ids, which can complicate having unique station names down the line - e.g. additional sites on the same stream or another agency monitoring the same stream.
* In Luca's siteinfo worksheet, he mentions some sites that may not be useful for our thermal sensitivity analysis. Flag those here.
* Add date-time and year for plotting.

```{r add new fields, message=FALSE, warning=FALSE}

fsdat <- left_join(fsdat, usfs_dpmd %>%
                     select(data_SiteID, Agency_ID),
                   by = c("file_name" = "data_SiteID")) 
fsdat
```

Combined new data for four Kenai sites with larger dataset.

```{r Combine Data Sets}

fsdat <- kendat %>%
  rename( sampleTime = Time,
  sampleDate = Date) %>% 
  bind_rows(fsdat)
fsdat
```
# Drop sites here
Drop sites highlighted in yellow on worksheet and any additional sites covered in conversation with Luca A.

Worksheet drops  include: 
"Clear Creek", "Hatchery Creek", "Ibeck Creek", "Jack Bay River", "McKinley Lake", "NF Williwaw Creek", "Pigot Bay Spawn Channel", "Rude River SC",
"SF Williwaw Creek", "Solf Lake Inlet" &"Steller Jay Creek"

Additional drops: Eyak Lake Trib

Possible: Eagle Creek and Shelter Bay are tidally influenced but will keep for now

```{r}
fschdat <- tibble()
fschdat <- fsdat %>% 
  mutate(useSite = case_when(
    Agency_ID %in% c("Clear Creek", "Hatchery Creek", "Ibeck Creek",
                     "Jack Bay River", "McKinley Lake", "NF Williwaw Creek",
                     "Pigot Bay Spawn Channel", "Rude River SC",
                     "SF Williwaw Creek", "Solf Lake Inlet",
                     "Steller Jay Creek") ~ 0, TRUE ~ 1),
    dt = as.POSIXct(paste(sampleDate, sampleTime, sep = " "),
                    format = "%Y-%m-%d %H:%M"),
    SiteID = Agency_ID,
    year = year(sampleDate)) %>% 
  select(-Date, -file_name, -year1)

fschdat
```

# STOPPED HERE - ADDED IN NEW DATA 1/25/21



Quick plots of data to make sure they read in ok. Start by summarizing daily means because quicker to plot. Looks like some bad winter temps well below zero that could be clipped later. No obvious air temps in summer as everything is < = 20 or so.

```{r plot of dialy means, message=FALSE, warning=FALSE}

fschdat %>% 
  group_by(SiteID, sampleDate) %>% 
  summarize(meant = mean(Temperature)) %>% 
  ggplot(aes(x = sampleDate, y = meant)) +
  geom_line() +
  facet_wrap(~SiteID)

```

Wrong! The sub-daily temps show more errors, definitely some air temps that need to be removed. This will be a good dataset for testing scripts, although it will probably need cleaning sooner rather than later for AKSSF.

```{r plot of raw data}

fschdat %>% 
  ggplot(aes(x = dt, y = Temperature)) +
  geom_line() +
  facet_wrap(~SiteID)


```

Rolling pdf of raw data to send to Luca and check on status of data QA.

```{r plot of raw data by site-year}
fs_sites <- fschdat %>% distinct(Agency_ID, year) %>% arrange(Agency_ID, year)

pdf("data_preparation/USFS Raw Data by Site and Year.pdf", width = 11, height = 8.5)

for(i in 1:nrow(fs_sites)) {
  dat <- left_join(fs_sites %>% slice(i), fschdat)
  subtitle <- dat %>% distinct(useSite) %>% pull(useSite)
  p1 <- dat %>% 
    ggplot(aes(x = dt, y = Temperature)) +
    geom_line() +
    labs(title = fs_sites %>% slice(i) %>% unite(site_year) %>%
           pull(site_year),
         subtitle = paste0("Use Site: ", subtitle)) +
    theme(legend.position = "bottom")
  print(p1)
}

dev.off()

```


Save file for summary report.

```{r}
fschdat

names(fschdat)

fschdat %>% 
  rename(SiteID = Agency_ID) %>% 
  select(SiteID, useSite, sampleDate, sampleTime, dt, year, Temperature) %>% 
  saveRDS("output/fschdat.rds")
```


