---
title: "data_kod_Tribes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE, messages = FALSE)
knitr::opts_knit$set(root.dir = normalizePath("..")) #this sets the root.dir up one level back to the project so that paths are relative to the project directory.

library(readxl)
library(xlsx)
library(stringr)
library(lubridate)
library(googlesheets4)
library(rnoaa)
library(hms)
library(tidyverse)
library(googledrive) 
```


```{r Functions}

# Source functions in helper function script to ensure proper format of saved data
source("helper_functions.R")

acronym = "kod_Tribes_"

```
# AKOATS
DM Notes: Use GOOGLE SHEET Copy for consistency

```{r AKOATS MD}

akoats_2020wk <- "https://docs.google.com/spreadsheets/d/1SPZXNGm_Tc39-GuJXY8j7Eb1lX6DwXTQ1en2_LvCI1I/edit#gid=1281874712"

akoats_kod_Tribes <- read_sheet(akoats_2020wk, sheet = "AKOATS_COMPLETE",
                    col_names = TRUE,
                    col_types = "c") %>% 
  # select(seq_id,Agency_ID,Contact_person,SourceName,Contact_email,
  #        Contact_telephone,Latitude,Longitude,Sensor_accuracy) %>%
  rename(AKOATS_ID = seq_id) %>% 
         #SiteID = Agency_ID) %>% 
  mutate(AKOATS_ID = as.numeric(AKOATS_ID),
         Latitude = as.numeric(Latitude),
         Longitude = as.numeric(Longitude))

akoats_kod_Tribes <- akoats_kod_Tribes %>% 
  filter(SourceName == "TribAlutiiqOldHarbor"|SourceName == "TribLarsenBay")
akoats_kod_Tribes


```

# Read in data and format
Two files from Kodiak Tribes

```{r Get temp data}
getwd()
# Folder containing source datasets
source_fol <- ".\\data_preparation\\source\\16_Kodiak_Tribes\\source"

files <- list.files(source_fol, full.names = T, recursive = T, pattern = ".*.csv|.*.xlsx")

# pattern used to name air temp files
pattern <- "_at_"

# Remove airtemp files
files <- files[!grepl(pattern, files)]
print(basename(files))

#tibble for all data
kod_Tribes.data <- tibble()

#dt temp tibble
kod_Tribes.data.a <- tibble()

#date time temp tibble
kod_Tribes.data.b <- tibble()

for ( i in files){
  filename <-  as.character( basename( i))
  print( filename)
  dat <- read_excel( path = i, col_names = FALSE, skip = 10)
  # print( paste0( sapply( dat, class)))
  
  # Format for Datetime and Temp
  if ( class( dat$"...1") == "POSIXct" && class(dat$...2) == "numeric"){
    print( "Column 1 = dt|Column 2 = Temp")
    dat <- dat[1:2]
    colnames(dat) <- c("dt", "Temperature")
    dat <- dat %>% 
      mutate( data_filename = filename,
              sampleDate = as_date(dt),
              sampleTime = as_hms(dt),
              year = year(dt),
              UseData = 1,
              UseSite = 1)
    kod_Tribes.data.a <- bind_rows(kod_Tribes.data.a, dat[!is.na(dat$Temperature),])
    
    # Format for Date, Time and Temp  
  }else if (class(dat$"...1") == "POSIXct" && class(dat$"...2") == "POSIXct"){
    print( "Column 1 = Date | Column 2 = Time | Column 3  = Temp")
    dat <- dat[1:3]
    colnames(dat) <- c("sampleDate", "sampleTime", "Temperature")
    dat <- dat %>% 
    mutate( data_filename = filename,
            sampleTime = as_hms(sampleTime),
            sampleDate = as_date(sampleDate),
            dt = as.POSIXct(paste(sampleDate, sampleTime, sep = " "),
                    format = "%Y-%m-%d %H:%M", tz = "GMT"),
            year = year(dt),
            UseData = 1,
            UseSite = 1)
    
    kod_Tribes.data.b <- bind_rows(kod_Tribes.data.b, dat[!is.na(dat$Temperature),])
    
    # Format for Plot Number, DateTime and Temp    
  }else if (class(dat$"...2") == "POSIXct" && class(dat$"...3") == "numeric"){
    print( "Column 1 = Meas# | Column 2 = DateTime | Column 3 = Temp")
    dat <- dat[2:3]
    colnames(dat) <- c("dt", "Temperature")
    dat <- dat %>% 
      mutate( data_filename = filename,
              sampleDate = as_date(dt),
              sampleTime = as_hms(dt),
              year = year(dt),
              UseData = 1,
              UseSite = 1)
    
    kod_Tribes.data.a <- bind_rows(kod_Tribes.data.a, dat[!is.na(dat$Temperature),])
    
    # Format for Plot Number, Date, Time and Temp    
  }else if (class(dat$"...2") == "POSIXct" && class(dat$"...3") == "POSIXct"){
    print( "Column 1 = Meas# | Column 2 = Date | Column 3 = Time | Column 4 = Temp" )
    dat <- dat[2:4]
    colnames(dat) <- c("sampleDate", "sampleTime", "Temperature")
    dat <- dat %>% 
    mutate( data_filename = filename,
            sampleTime = as_hms(sampleTime),
            sampleDate = as_date(sampleDate),
            dt = as.POSIXct(paste(sampleDate, sampleTime, sep = " "),
                    format = "%Y-%m-%d %H:%M", tz = "GMT"),
            year = year(dt),
            UseData = 1,
            UseSite = 1)
    
    kod_Tribes.data.b <- bind_rows(kod_Tribes.data.b, dat[!is.na(dat$Temperature),])
    
  }else{
    print( paste0( "Format not recognized for file ", filename))
  }
  
}

kod_Tribes.data <- bind_rows(kod_Tribes.data.a, kod_Tribes.data.b)

dput(unique(kod_Tribes.data$data_filename))

# Create Temporary Site IDs for comparison and add Agency_ID from AKOATS
siteid <- tibble( 
  data_filename = c("kdk_bigcr01_10709379_20161104qc.xlsx",
                    "kdk_bigcr01a_20151019qc.csv.xlsx", "kdk_bigcr01a_20161104qc.xlsx",
                    "kdk_bigcr01a_20180801qc.xlsx", "kdk_bigcr01a_20181024qc.xlsx",
                    "kdk_bigcr01b_20151019qc.csv.xlsx", "kdk_bigcr01b_20161104qc.xlsx",
                    "kdk_karrv_02a_20151030qc.xlsx", "kdk_karrv_02a_20170712qc.xlsx",
                    "kdk_karrv_02a_20180714qc.xlsx", "kdk_karrv_02a_20181108qc.xlsx",
                    "kdk_karrv_02b_20151030qc.xlsx", "kdk_karrv_02b_20170712qc.xlsx",
                    "kdk_karrv_02b_20180714qc.xlsx", "kdk_karrv_02c_20180714qc.xlsx",
                    "kdk_karrv_02c_20181108qc.xlsx", "kdk_bigcr01a_20160405qc.xlsx",
                    "kdk_bigcr01b_20160629qc.xlsx", "kdk_karrv_02a_20160903qc.xlsx",
                    "kdk_karrv_02a_20171020qc.xlsx", "kdk_karrv_02b_20160903qc.xlsx",
                    "kdk_karrv_02b_20171020qc.xlsx", "kdk_karrv_02c_20171020qc.xlsx"),
  
  SiteID_temp = c("big creek",
                  "big creeka", "big creeka",
                  "big creeka", "big creeka",
                  "big creekb", "big creekb",
                  "karluka", "karluka",
                  "karluka", "karluka",
                  "karlukb", "karlukb",
                  "karlukb", "karlukc",
                  "karlukc", "big creeka",
                  "big creekb", "karluka",
                  "karluka", "karlukb",
                  "karlukb", "karlukc"),
  
  Agency_ID = c("kdk_bigcr01",
                  "kdk_bigcr01", "kdk_bigcr01",
                  "kdk_bigcr01", "kdk_bigcr01",
                  "kdk_bigcr01", "kdk_bigcr01",
                  "kdk_karrv02", "kdk_karrv02",
                  "kdk_karrv02", "kdk_karrv02",
                  "kdk_karrv02", "kdk_karrv02",
                  "kdk_karrv02", "kdk_karrv02",
                  "kdk_karrv02", "kdk_bigcr01",
                  "kdk_bigcr01", "kdk_karrv02",
                  "kdk_karrv02", "kdk_karrv02",
                  "kdk_karrv02", "kdk_karrv02"))

kod_Tribes.data <- kod_Tribes.data %>% 
  left_join(siteid)

kod_Tribes.data %>% 
 count(SiteID_temp, year)
  

```

## Create Metadata
Pull Site info from AKOATS
Seems to be duplicate loggers at each site (a,b,c suffixes) - plot by dataset first and compare

```{r Save Metadata}

# akoats_fields <- c ("SiteID", "seq_id", "Agency_ID", "SourceName", "Contact_person", 
#                      "Contact_email", "Contact_telephone", "Latitude", "Longitude", 
#                      "Sensor_Placement", "Waterbody_name", "Waterbody_type", "Sensor_accuracy", 
#                      "Sensor_QAQC")
# kod_Tribesmd <- kod_Tribesmd %>% 
#   select(all_of(akoats_fields))
# 
# kod_Tribesmd
# 
# # Save Metadata
# save_metadata_files(kod_Tribesmd, acronym)

```
## Rolling pdf by site and year
Split by waterbody because there are multiple loggers/site
No data for big creek during summertime in 2017

```{r plot of raw data by site-year}
getwd()

kod_Tribes.data.qc <- kod_Tribes.data %>% 
  filter( month(sampleDate) %in% 6:9)

kod_Tribes_sites <- kod_Tribes.data.qc %>% distinct(Agency_ID, year) %>% arrange(Agency_ID, year)

pdf("./data_preparation/kod_Tribes Raw Data by Site and Year.pdf", width = 11, height = 8.5)
# Get limits of temp data
for(i in 1:nrow(kod_Tribes_sites)) {
  dat <- left_join(kod_Tribes_sites %>% slice(i), kod_Tribes.data.qc)
  subtitle <- dat %>% distinct(Agency_ID) %>% pull(Agency_ID)
  xmin <- as.POSIXct(min(dat$dt),format = "%Y-%m-%d %H:%M")
  xmax <- as.POSIXct(max(dat$dt),format = "%Y-%m-%d %H:%M")
  p1 <- dat %>%
    ggplot(aes(x = dt, y = Temperature)) +
    geom_line(aes(color = SiteID_temp, linetype = SiteID_temp)) +
    scale_x_datetime(limits = c(xmin, xmax), labels = waiver()) +
    scale_y_continuous(limits = c(-5, 30), labels = waiver()) +
    labs(title = kod_Tribes_sites %>% slice(i) %>% unite(agencyid_year) %>%
           pull(agencyid_year),
         subtitle = paste0("Agency_ID: ", subtitle)) +
    theme(legend.position = "bottom")
  print(p1)
}

dev.off()

```
# Interactive Plot
Examine data to identify loggers with most complete coverage  for time period of interest and remove obvious errors

```{r Interactive Plot With Air Temp and Tide}

# Filter by site: kdk_bigcr01 & kdk_karrv02
p <- kod_Tribes.data.qc %>% 
  filter(Agency_ID == "kdk_karrv02")

xmin <- as.POSIXct(min(p$dt),format = "%Y-%m-%d %H:%M", tz = "GMT")
xmax <- as.POSIXct(max(p$dt),format = "%Y-%m-%d %H:%M", tz = "GMT")

p <- p %>% 
  ggplot() +
  #geom_line( data = cd.climDat, aes(x = dt, y = tmin, color = "Air min")) +
  #geom_line( data = cd.climDat,aes(x = dt, y = tmax, color = "Air max")) +
  #geom_line( data = cd.tideDat,aes(x = dt, y = verified_meters, color = "Tide (m)")) +
  geom_line(aes(x = dt, y = Temperature, color = SiteID_temp, linetype = SiteID_temp)) +
  scale_x_datetime(limits = c(xmin, xmax), labels = waiver()) +
  coord_cartesian(ylim = c(-5, 30)) + 
  facet_wrap(~Agency_ID) +
  theme(legend.title = element_blank()) +
  labs(title = "Temperature by Site",
       y = "Temperature degrees C",
       x = "Time of Measurement")

ggplotly(p)

```

# Filter out extra logger and average dups
Both Big Creek and Karluk River have duplicate loggers.  Big Creek should have measurements from Bigcreeka and Bigcreekb averaged and Karluk should have measurements from Karluka and Karlukb averaged.  I only looked at measurements for the time period of interest so these loggers may not provide the most complete series for the entire range but are the best fit for the AKSSF project

```{r Average temp measurements and save}

# Prepare dataset for QC
kod_Tribes.data.final <- kod_Tribes.data %>% 
  filter( SiteID_temp %in% c("karluka","karlukb","big creeka", "big creekb"),
          ( month(sampleDate) %in% 6:9)) %>% 
  select( Agency_ID, sampleDate, sampleTime, Temperature, UseData, UseSite,
                        year, dt) %>%
  group_by(Agency_ID, dt) %>% 
  mutate( avgTemp = mean(Temperature, na.rm = TRUE)) %>% 
  ungroup %>% 
  mutate(Temperature = avgTemp,
         SiteID = Agency_ID) %>% 
  select(-avgTemp) %>% 
  distinct()

#Reorder and save output
colorder <- c( "SiteID","Agency_ID", "sampleDate", "sampleTime",
              "Temperature", "dt", "year", "UseSite", "UseData")

kod_Tribes.data.final <- kod_Tribes.data.final[,colorder]

kod_Tribes.data.final %>% 
  count(Agency_ID, year)


# Save copy of the averaged temp data for QC

saveRDS(kod_Tribes.data.final, "./data_preparation/formatted_data/kod_Tribes.data.rds")
```

# Plot and Rolling PDF of cleaned data
Plot of cleaned/averaged temp data ready for qc....Data may have already been qd'd by Jeff Davis, awaiting email response

```{r plot of raw data by site-year}
getwd()

kod_Tribes_sites <- kod_Tribes.data.final %>% distinct(Agency_ID, year) %>% arrange(Agency_ID, year)

pdf("./data_preparation/kod_Tribes Cleaned Data by Site and Year.pdf", width = 11, height = 8.5)
# Get limits of temp data
for(i in 1:nrow(kod_Tribes_sites)) {
  dat <- left_join(kod_Tribes_sites %>% slice(i), kod_Tribes.data.final)
  subtitle <- dat %>% distinct(Agency_ID) %>% pull(Agency_ID)
  xmin <- as.POSIXct(min(dat$dt),format = "%Y-%m-%d %H:%M")
  xmax <- as.POSIXct(max(dat$dt),format = "%Y-%m-%d %H:%M")
  p1 <- dat %>%
    ggplot(aes(x = dt, y = Temperature)) +
    geom_line() +
    scale_x_datetime(limits = c(xmin, xmax), labels = waiver()) +
    scale_y_continuous(limits = c(-5, 30), labels = waiver()) +
    labs(title = kod_Tribes_sites %>% slice(i) %>% unite(agencyid_year) %>%
           pull(agencyid_year),
         subtitle = paste0("Agency_ID: ", subtitle)) +
    theme(legend.position = "bottom")
  print(p1)
}

dev.off()

```
