---
title: "data_CIK"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath(".."))

library(broom)
library(readxl)
library(stringr)
library(lubridate)
library(googlesheets4)
library(rnoaa)
library(hms)
library(tidyverse)


```


# Data archived on KNB - sites through 2017.

Read in data files saved on S drive. These are all the data that Sue archived on KNB. These include some Bristol Bay data that will have to be removed. The site IDs for just Cook Inlet sites all have "CIK" in the name.

```{r}
cik.files <- list.files("S:/Stream Temperature Data/Cook Inletkeeper", pattern = ".csv", full.names = TRUE)

cik.files <- cik.files[!grepl("SiteLevelMetadata_Mauger", cik.files)]

```

Metadata

```{r}
cik_md <- read_csv("S:\\Stream Temperature Data\\Cook Inletkeeper\\SiteLevelMetadata_Mauger.csv") 

cik_knb_md <- cik_md %>% 
  filter(grepl("CIK", SiteID)) %>% 
  select(AKOATS_ID, SiteID, SourceName, Contact_person, Latitude, Longitude, Waterbody_name)

cik_knb_md

```


Combined data files and get information from metadata file.

```{r}
cik.data <- cik.files %>% 
  map_df(function(x) read_csv(x) %>% 
  mutate(file_name = gsub(".csv","",basename(x))))

cik.data <- left_join(cik.data, cik_knb_md %>% select(AKOATS_ID, SiteID, Waterbody_name)) %>% 
  filter(grepl("CIK", SiteID))

cik.data #removed ~500K rows of data that were from Bristol bay, still > 4M rows of data for cook inlet.
```

# New data

Sue sent over two zip files saved in separate folders. One is a submission to NCEAS that she thinks didn't get uploaded -- folder name is "new submissions Dec 2018". It looks like some data from 2017 to 2018. The second zip folder has data from 2018-2020 and is called "UAA". There is a metadata file in each one with site IDs that should link to AKOATs.

I'll remove the Bristol Bay sites and read them in separately in the SWSHP repo. All the Bristol Bay data are being used for one project with Erik and then also being brought into this repo for the larger AKSSF project.

```{r}
cik_folder <- ("S:/Stream Temperature Data/Cook Inletkeeper")

cik_files1 <- list.files(paste(cik_folder, "/February_2021/new submissions Dec 2018", sep = ""), full.names = TRUE)
cik_files1 <- cik_files1[!grepl("new submissions Dec 2018.xlsx", cik_files1)]

cik_files2 <- list.files(paste(cik_folder, "/February_2021/UAA", sep = ""), full.names = TRUE)
cik_files2 <- cik_files2[!grepl("new submissions Jan 2021.xlsx", cik_files2)]

```

Read in each set of data and combine them.

```{r}

cik_dat1 <- map_df(cik_files1, function(x) read_excel(x) %>% mutate(file_name = gsub(".xlsx", "", basename(x))))

summary(cik_dat1)

#two time fields are independent
cik_dat1 %>% 
  count(is.na(Time), is.na(`Time, GMT-08:00`))

#for site naming pattern, trim last 8 characters (_xx_xxxx) to get SiteID
cik_dat1 %>% distinct(file_name)

cik_dat1 <- cik_dat1 %>% 
  mutate(sampleDate = as.Date(Date),
         sampleTime = case_when(is.na(`Time, GMT-08:00`) ~ as_hms(Time),
                                TRUE ~ as_hms(`Time, GMT-08:00`)),
         Temperature = `Temp, °C`,
         SiteID = substr(file_name, 1, nchar(file_name) - 8))  %>% 
  select(SiteID, sampleDate, sampleTime, Temperature)
  
```

```{r}
cik_dat2 <- map_df(cik_files2, function(x) read_excel(x) %>% mutate(file_name = gsub(".xlsx", "", basename(x))))

summary(cik_dat2)

#for site naming pattern, trim last 8 characters (_xx_xxxx) to get SiteID
cik_dat2 %>% distinct(file_name)

cik_dat2 <- cik_dat2 %>% 
  mutate(sampleDate = as.Date(Date),
         sampleTime = as_hms(`Time, GMT-08:00`),
         Temperature = `Temp, °C`,
         SiteID = substr(file_name, 1, nchar(file_name) - 8))  %>% 
  select(SiteID, sampleDate, sampleTime, Temperature)
  
cik_dat <- bind_rows(cik_dat1, cik_dat2) %>% 
  filter(grepl("CIK", SiteID))

cik_dat %>% distinct(SiteID)
```

Make sure that SiteIDs in the data link to AKOATs and get the waterbody name so that we have it for plotting.

```{r}
akoats <- read_excel("S:/EPA AKTEMP/AKOATS_DATA_2020_working.xlsx", sheet = "AKOATS_COMPLETE") 

#filter on continuous sites
akoats <- akoats %>% 
  filter(Sample_interval == "continuous")

```

All are there.

```{r}
cik_dat <- cik_dat %>% 
  left_join(akoats %>% select(Agency_ID, seq_id, Waterbody_name), by = c("SiteID" = "Agency_ID")) %>% 
  rename(AKOATS_ID = seq_id)

```

Combine with 2017 data.

```{r}
intersect(names(cik.data), names(cik_dat))
cik_dat <- bind_rows(cik.data, cik_dat)

cik_dat %>% distinct(SiteID, AKOATS_ID, Waterbody_name)
```

# Review data

Add year and date time field for plotting.

```{r}
cik_dat <- cik_dat %>% 
  mutate(year = year(sampleDate),
         dt = parse_datetime(paste(sampleDate, sampleTime), format = "%Y-%m-%d %H:%M:%S")) 
```

Plot of raw data, but Sue typically provides only QAed data. Some useData==0

```{r}
cik_dat %>% count(UseData)

cikSites <- cik_dat %>% distinct(SiteID, year)

pdf("output/CIK raw data by site.pdf")

for(i in 1:nrow(cikSites)) {
  dat <- left_join(cikSites %>% slice(i), cik_dat)
  p1 <- dat %>% 
    filter(!UseData == 0) %>%
    ggplot(aes(x = dt, y = Temperature)) +
    geom_line() +
    facet_wrap(~year) +
    labs(title = dat %>% distinct(SiteID))
  print(p1)  
}

dev.off()

```

Number of years of data for each site.

```{r}
cikSites %>% 
  count(SiteID)

ltSites <- cikSites %>% 
  count(SiteID) %>% 
  filter(n > 10) %>% 
  pull(SiteID)
```

Plot of monthly temperatures.

```{r}
cik_dat %>% 
  filter(SiteID %in% ltSites) %>% 
  group_by(SiteID, Waterbody_name, year, month = month(sampleDate)) %>% 
  summarize(monMean = mean(Temperature)) %>% 
  ggplot(aes(x = year, y = monMean, color = as.factor(month))) + 
  geom_point() +
  # geom_smooth(method = "lm") +
  facet_wrap(~Waterbody_name)
```

Correlation between monthly temperatures and year by Site.

```{r}
cik_monthly <- cik_dat %>% 
  filter(SiteID %in% ltSites) %>% 
  group_by(SiteID, Waterbody_name, year, month = month(sampleDate)) %>% 
  summarize(monMean = mean(Temperature)) 
  
cik_monthly %>% 
  group_by(SiteID, month) %>% 
  summarize(correlation = cor(year, monMean, use = "complete.obs")) %>% 
  pivot_wider(names_from = SiteID, values_from = correlation)

```
```{r}
monthlyLMs <- cik_monthly %>% 
  nest(data = c(-Waterbody_name, -month)) %>% 
  mutate(
    fit = map(data, ~ lm(monMean ~ year, data = .x)),
    tidied = map(fit, tidy)
  ) %>% 
  unnest(tidied)

sig_trends <- monthlyLMs %>% 
  filter(p.value < 0.05, term == "year") %>% 
  select(Waterbody_name, month)

monthlyLMs %>% 
  filter(p.value < 0.05, term == "year") %>% 
  ggplot(aes(x = as.factor(month), y = estimate, color = Waterbody_name)) +
  geom_point() +
  theme_minimal() +
  geom_abline(aes(intercept = 0, slope = 0)) 

```


Table of sample sizes.

```{r}
cik_monthly %>% 
  right_join(sig_trends) %>% 
  group_by(SiteID, Waterbody_name, month) %>% 
  summarize(n = n())

```
Those estimates are really high, ~0.2 deg per year. Plot just these monthly trends and see if average monthly temperatures have actually increasd by 2 deg over the last decade.

```{r}
cik_monthly %>% 
  right_join(sig_trends) %>% 
  ggplot(aes(x = year, y = monMean, color = as.factor(month))) + 
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~Waterbody_name) +
  theme_minimal()
```

Would we see more significant results if we clipped the other sites to 2005 or later? Yes, 34 sig. trends in monthly means over last 15 years as compared to 24 if we don't clip the data.

```{r}
monthlyLMs2 <- cik_monthly %>% 
  filter(year > 2005) %>% 
  nest(data = c(-Waterbody_name, -month)) %>% 
  mutate(
    fit = map(data, ~ lm(monMean ~ year, data = .x)),
    tidied = map(fit, tidy)
  ) %>% 
  unnest(tidied)

sig_trends2 <- monthlyLMs2 %>% 
  filter(p.value < 0.05, term == "year") %>% 
  select(Waterbody_name, month)

nrow(sig_trends2) - nrow(sig_trends)

monthlyLMs2 %>% 
  filter(p.value < 0.05, term == "year") %>% 
  ggplot(aes(x = as.factor(month), y = estimate, color = Waterbody_name)) +
  geom_point() +
  theme_minimal() +
  geom_abline(aes(intercept = 0, slope = 0)) +
  labs(x = "Month", y = "Annual Trend in Mean Monthly Temperature", title = "Significant Trends (p-value < 0.05) in Mean Monthly Temperatures \n from 2005-2019 for Cook Inlet Streams", color = "Stream: ") +
  theme(legend.position = "bottom")

ggsave("output/CIK monthly trends.pdf", width = 10, height = 8)
```


