---
title: "data_ADFG"
author: "dwmerrigan"
date: "3/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE, messages = FALSE)
knitr::opts_knit$set(root.dir = normalizePath("..")) #this sets the root.dir up one level back to the project so that paths are relative to the project directory.

library(readxl)
library(xlsx)
library(stringr)
library(lubridate)
library(googlesheets4)
library(rnoaa)
library(hms)
library(tidyverse)
library(googledrive) 

```

```{r Functions}

# Source functions in helper function script to ensure proper format of saved data
source("helper_functions.R")

acronym = "adfgKod_"

```


# Read in data and format
6 Kodiak sites from Heather Finkle 

Karluk River: Tides >= 9' can influece water flow @ Karluk

Ayakulik River:
2019 have shuttle data incorrectly downloaded
River is subject to high water events

Buskin River:
2015 was the first season of using the temperature data loggers.
2020: there were issues with the shuttles and loggers not downloading the data or not launching correctly, which caused data gaps.
Buskin River is very susceptible to flash flooding and drought. It is also a very popular river for sport fishing and the loggers have been moved  or stolen by the public in the past.

Dog Salmon Sites:
Loggers deployed in 2015. 
Logger A is located in a westerly fork in the river by the weir and logger B is in a more easterly fork by the weir.
Both Dog Salmon Creek loggers were removed from the water in mid March, 2018. These loggers were replaced May 23, 2018 after field crews returned to camp.

Olga Creek:
2016 was the first season of using the temperature data loggers.
The location where the temperature logger is located is tidally influenced.

```{r Get temp data}
getwd()
# Folder containing source datasets
source_fol <- ".\\data_preparation\\source\\14_ADFG_Heather_Finkle\\source"

files <- list.files(source_fol, full.names = T, recursive = T, pattern = ".*.csv|.*.xlsx")
basename(files)
# Remove metadata sheets
patterns <- c("AKOATS")

akoats_md <- files
adfg_files <- files
for (pattern in patterns){
  adfg_files <- adfg_files[!grepl(pattern, adfg_files)]
  akoats_md <- akoats_md[grepl(pattern, akoats_md)]
}

basename(adfg_files)
adfg_dat <- tibble()

for (i in adfg_files) {
  filename <- basename(i)
  # Dog salmon has two sites in excel sheet
  if (grepl("Dog SalmonWater", filename)) {
    dat_a <-  read_excel(path = i, sheet = 2)
    dat_b <-  read_excel(path = i, sheet = 3)
    dat_a <-
      dat_a[2:3] %>% mutate(data_filename = paste0(filename, "a"))
    dat_b <-
      dat_b[2:3] %>% mutate(data_filename = paste0(filename, "b"))
    colnames(dat_a) <- c("dt", "Temperature", "data_filename")
    colnames(dat_b) <- c("dt", "Temperature", "data_filename")
    # Format data
    dat <- bind_rows(dat_a, dat_b) %>%
      transform(
        sampleDate = as_date(dt),
        sampleTime = as_hms(dt),
        year = year(dt),
        Temperature = as.numeric(Temperature)
      )
    adfg_dat <-
      bind_rows(adfg_dat, dat[!is.na(dat$Temperature),]) %>%
      select(data_filename,
             dt,
             sampleDate,
             sampleTime,
             year,
             Temperature)
    
  } else if(!grepl("Dog SalmonWater", filename)) {
    filename <- basename(i)
    dat <-  read_excel(path = i, sheet = 2)
    dat <- dat[2:3]
    colnames(dat) <- c("dt", "Temperature")
    print(colnames(dat))
    
    # Format data
    dat <- dat %>% mutate(data_filename = filename) %>%
      transform(
        sampleDate = as_date(dt),
        sampleTime = as_hms(dt),
        year = year(dt),
        Temperature = as.numeric(Temperature)
      )
    adfg_dat <-
      bind_rows(adfg_dat, dat[!is.na(dat$Temperature),]) %>%
      select(data_filename,
             dt,
             sampleDate,
             sampleTime,
             year,
             Temperature)
  }
}


```

```{r Create table to link Agency ID to filename}
adfgmd <- tibble( data_filename = c("AyakulikR_watertemp2016-2020.xlsx",
                           "BuskinRiver Water Temp2015-2020.xlsx",
                           "Dog SalmonWater Temp2015-2020.xlsxa",
                           "Dog SalmonWater Temp2015-2020.xlsxb",
                           "KarlukWater Temp2015-2020.xlsx",
                           "Olga Creek Water Temp2016-2020.xlsx"),
                  
          Agency_ID = c("kdk_ayarv01","kdk_busrv01","kdk_doscr01",
                        "kdk_doscr02", "kdk_karrv01", "kdk_olgcr01a"))

akoats_md <- list.files(source_fol, full.names = T, recursive = T, pattern = "AKOATS")

akoats_kdkmd <- akoats_md %>% 
  read_excel() %>% 
    left_join(adfgmd)

akoats_kdkmd

```

```{r Join AGENCY_ID to temp data}

adfgKod.data <- adfg_dat %>% 
  left_join(akoats_kdkmd, by = "data_filename") %>% 
  rename(SiteID=Agency_ID) %>% 
  mutate(UseData = 1, UseSite = 1) %>% 
  select(SiteID, sampleDate, sampleTime, dt, year, Temperature, UseData, UseSite) %>% 
  distinct()

summary(adfgKod.data)

```

# Rolling pdf of sites by year
```{r plot of raw data by site-year}
# Filter on period of interest
adfgKod.data.qc <- adfgKod.data %>% 
  filter( month(sampleDate) %in% 6:9)

saveRDS(adfgKod.data.qc, "data_preparation/formatted_data/adfgKod.data.rds")

adfgKod_sites <- adfgKod.data.qc %>% distinct(SiteID, year) %>% arrange(SiteID, year)

pdf("data_preparation/adfg Kodiak Raw Data by Site and Year.pdf", width = 11, height = 8.5)
# Get limits of temp data 
for(i in 1:nrow(adfgKod_sites)) {
  dat <- left_join(adfgKod_sites %>% slice(i), adfgKod.data.qc)
  subtitle <- dat %>% distinct(UseSite) %>% pull(UseSite)
  xmin <- as.POSIXct(min(dat$dt),format = "%Y-%m-%d %H:%M")
  xmax <- as.POSIXct(max(dat$dt),format = "%Y-%m-%d %H:%M")
  p1 <- dat %>% 
    ggplot(aes(x = dt, y = Temperature)) +
    geom_line() +
    scale_x_datetime(limits = c(xmin, xmax), labels = waiver()) +
    scale_y_continuous(limits = c(-5, 30), labels = waiver()) +
    labs(title = adfgKod_sites %>% slice(i) %>% unite(site_year) %>%
           pull(site_year),
         subtitle = paste0("Use Site: ", subtitle)) +
    theme(legend.position = "bottom")
  print(p1)
}

dev.off()

```
# Save metadata
## Create Metadata
One site not in AKOATS but data provider has included AKOATS md

```{r Save Metadata}

akoats_fields <- c ("SiteID", "seq_id", "Agency_ID", "SourceName", "Contact_person", 
                     "Contact_email", "Contact_telephone", "Latitude", "Longitude", 
                     "Sensor_Placement", "Waterbody_name", "Waterbody_type", "Sensor_accuracy", 
                     "Sensor_QAQC")

akoats_kdkmd <- akoats_kdkmd %>% 
  mutate(SiteID = Agency_ID) %>% 
  select(all_of(akoats_fields))

akoats_kdkmd

# Save Metadata
save_metadata_files(akoats_kdkmd, acronym)

```
